As we pointed out in~\ref{subsec:security-model-preliminaries.} the protocol requires a common notion of time among all participants.
To avoid relying on centralized time oracles which would undermine network security we adopt\
decentralized logical time synchronization technique based on synchronization beacons~\cite{cryptoeprint:2019/838}.

All committees $V_{n}^k$ members participate in the synchronization process,\
and it is based on the following logical blocks:
\begin{legal}
    \item[]\textbf{Synchronization slots.} Once a consensus participant’s local time reaches synchronization\
    slot ${sl_{n \cdot R}, n \in \mathbb{N}}$, his clocks are adjusted before moving to the\
    next slot (i.e.\ next epoch).
    \item[]\textbf{Synchronization beacons.} In addition to other messages, all members of the consensus group\
    generate so-called synchronization beacons.
    For every local slot $sl_{i, j}^{\text{loc}} \in [n \cdot R + 1,\dots,n \cdot R + R \// 6], n \in \mathbb{N}$\
    every $PK_i$ evaluates $\mathcal{F}_{\text{VRF}}$ functionality with input\
    ${x_{i, j}^{\textsf{sync}, n} = (\eta_n || sl_{i, j}^{\text{loc}})}$ to get a proof $\pi_{i, j}^{\text{sl}}$\
    and checks if he has the right to release a beacon by comparing the pseudo-random value\
    ${y_{i, j}^{\textsf{sync}} = \mathcal{H}(r_{i, j}^{\text{sl}} || \textsf{SYNC})}$\
    with the corresponding threshold $T_{i, n}^{\text{sync}} = 2^{l_{\text{VRF}}} \cdot \phi(\alpha_{i}^{n - 2})$.
    If ${y_{i, j}^{\textsf{sync}} < T_{i, n}^{\text{sync}}}$ then the participant broadcasts a beacon\
    $b_{i, j}^{\text{sync}} = (v^{\text{vrf}}_i, sl_{i, j}^{\text{loc}}, \pi_{i, j}^{\text{sl}})$.
    \item[]\textbf{Arrival times bookkeeping.} Every consensus participant $PK_i$ maintains\
    an array $\mathbf{b}_i^{\text{set}}$ of received beacons with beacon's\
    arrival local time ${sl^{\text{rec}}_{i, j}: (sl_{i', j}^{\text{loc}}, \text{flag})\
    \in \mathbb{N} \times (\text{final}, \text{temp})}$.
    Assume a beacon $b_{i', j'}^{\text{sync}}$ emitted by $PK_{i'}$ is fetched by a party $PK_i$ for the first time:
    \begin{itemize}
        \item If $PK_i$ has not yet passed synchronization slot $sl_{i, n \cdot R}^{\text{loc}}$ and the\
        received beacon belongs logically to this party’s next epoch, then decision is marked\
        as temporary and $PK_i$ stores a record\
        ${{sl^{\text{rec}}_{i, j}}: (sl_{i', j'}^{\text{loc}}, \text{temp})}$.
        Value ${{sl^{\text{rec}}_{i, j}}}$ will be adjusted once this party adjusts\
        its local time-stamp for the next epoch.
        \item If $PK_i$ has already passed synchronization slot $sl_{i, n \cdot R}^{\text{loc}}$ but not yet passed\
        slot $sl_{i, (n + 1) \cdot R}^{\text{loc}}$, then the received time is defined as the current local slot\
        number and is considered final, i.e.\
        ${{sl^{\text{rec}}_{i, j}}: (sl_{i', j'}^{\text{loc}}, \text{final})}$.
    \end{itemize}
    If a party has already received a beacon for the same slot $j'$ and creator $PK_{i'}$, it will set\
    the arrival time equal to the first one received among those.
    \item[]\textbf{The synchronization interval.} For a local clock adjustment, which is triggered by a synchronization\
    slot only beacons with recorded arrival time\
    in the interval ${[(n - 1) \cdot R + 1,\dots,(n-1) \cdot R + R \// 6]}$ are used.
    \item[]\textbf{Computing the adjustment evidence.} The adjustment is computed based on the\
    received beacons set $\mathbf{b}_i^{\text{set}}$.
    Beacon ${b^{\text{rec}}_{i, j}}$ is only considered valid for adjusting procedure\
    triggered by a synchronization slot if:
    \begin{enumerate}
        \item Recorded time ${sl_{i', j'}^{\text{loc}} \in b_{i', j'}^{\text{sync}}}$ is final and belongs\
        to the synchronization interval ${[(n - 1) \cdot R + 1,\dots,(n-1) \cdot R + R \// 6]}$.
        \item Beacon is included into the block whose creation slot belongs\
        to the interval ${[(n - 1) \cdot R + 1, (n - 1) \cdot R + 2 \cdot R \// 3]}$
        \item Beacon's proof $\pi_{i, j}^{\text{sl}, n}$ is valid.
    \end{enumerate}
    \item[]\textbf{Adjusting the local clock.} Every party $PK_i$ computes $\textsf{shift}_{i, n}$ to\
    adjust its clock in the synchronization slot $sl_{n \cdot R}$.
    Value of the shift is calculated as\
    ${\hat{\textsf{shift}_{i, n}} = \textsf{median} \{sl_{i', j'}^{\text{loc}} - sl^{\text{rec}}_{i, j}\
    | \forall b_{i', j'}^{\text{sync}, n} \in B_i^{\text{sync}, n}\}}$.
    Using the calculated ${\longh{\textsf{shift}_{i, n}}}$ value all $\mathbf{b}_i^{\text{set}}$ records with\
    $\text{temp}$ status is updated as ${sl^{\text{rec}, n}_{i, j} + \hat{\textsf{shift}_{i, n}}}$ and\
    the adjustment procedure is performed again to receive a final $\text{shift}_{i, n}$ value.
\end{legal}


As a result of the above synchronization procedure, at the beginning of the next round the\
party $PK_i$ will report a local time equal to $n \cdot R + \text{shift}_{i, n} + 1$.
If ${\text{shift}_{i, n} > 0}$, the party proceeds by emulating its actions for shift rounds.
If ${\text{shift}_{i, n} < 0}$, the party remains a silent observer\
until its local time has advanced to slot $n \cdot R + 1$ and resumes normally at that round.